---
title: "Teaching & Mentorship"
# subtitle: ""
title-block-banner: false
toc: true
toc-location: left
page-layout: full
editor: 
  markdown: 
    wrap: 72
---

## Instruction

<b>Courses that I designed and taught</b><br>

### COMM 6200 Climate Communication and Community: Youth-Centered Participatory Action Research

::: grid
<!-- ::: {.g-col-12 .g-col-md-2} -->

<!-- ::: -->

::: {.g-col-12 .g-col-md-12}
[Spring 2025](https://www.asc.upenn.edu/graduate/courses#5219)

This doctoral-level, academically based community service (ABCS)
research seminar focuses on co-developing research questions with local
youth in West Philadelphia to explore their experiences with climate
change and identify and develop potentially implementable solutions for
how to mitigate, adapt to, and build resilience in the face of climate
change. The course involves the scheduled seminar, as well as required
fieldwork time at a partner high school. This course will apply
frameworks and theories from climate communication, behavior change, and
Youth Participatory Action Research (YPAR). Students will engage in a
hands-on, youth-engaged research project and apply YPAR principles to
empower youth voices and youth action in climate discourse and action.
Graduate students will learn core theories about behavior change,
communication intervention design, and YPAR. They will gain experience
designing and implementing a mixed-methods study, combining qualitative
with quantitative research techniques to conduct formative research,
message design, and testing in partnership with youth. Through this
project, students will develop proficiency in data analysis,
interpretation, and presentation of findings. The course will also cover
ethical and practical considerations in youth-centered research,
relationship building, community engagement strategies, and effective
facilitation skills. This course provides a unique opportunity for
doctoral students to gain practical experience in participatory research
while addressing pressing social and environmental issues in the West
Philadelphia community.
:::

:::

### Applied Research & Data Analysis Pro-seminar

::: grid
<!-- ::: {.g-col-12 .g-col-md-2} -->

<!-- ::: -->

::: {.g-col-12 .g-col-md-12}
[Spring 2021](https://dcosme.github.io/pltv-seminar/)

I designed an 8-week pro-seminar for University of Pennsylvania students
involved with [Penn Leads the Vote](https://www.pennvotes.org/) to get
hands on experience conducting applied research and analyzing data in
the context of their ongoing PLTV activities.

The pro-seminar was designed to teach students to:

-   Develop research questions related to their applied work
-   Identify ethical issues related to their research
-   Test research questions using observational and experimental methods
-   Design and implement surveys using Qualtrics
-   Execute and modify scripts to tidy, analyze, and visualize data in R
-   Summarize and interpret results
:::

:::

## Teaching assistance

<b>Courses for which I served as a Teaching Assistant</b><br>

::: {.grid}

::: {.g-col-12 .g-col-md-12}
**[Spring 2017 | PSY478/578 Social Development at the University of Oregon](https://pages.uoregon.edu/psychit/Syllabi_Archive/2010s/2017/Spring_2017/478_578-Pfeifer-Sp17.pdf)**
:::

:::

## Guest lectures

<b>Guest lectures I have given for undergraduate and graduate courses</b><br>

::: grid

::: {.g-col-12 .g-col-md-12}
Cosme, D. (2022) What motivates people to share content? Guest lecture
presented at the New York University PSYCH-UA 53 Psychological Science
and Society course, November 21.

Cosme, D. (2022) How can we help people make self-concordant decisions
to promote health and well-being? Guest lecture presented at the
University of Pennsylvania COMM 611 Neurobiology of Social Influence
course, October 5.

Cosme, D. (2022) [Specification Curve Analysis
Workshop](https://dcosme.github.io/specification-curves/SCA_tutorial_2022).
Guest lecture presented at the University of Pennsylvania COMM 783
Describing Your Data course, April 7.

Cosme, D. (2022) [Reproducibility
Workshop](https://dcosme.github.io/reproducibility-workshop/presentation).
Guest lecture presented at the University of Pennsylvania COMM 783
Describing Your Data course, March 17.

Cosme, D. (2021) Self development in adolescence. Guest lecture
presented at the University of Oregon PSY 631 Translational Neuroscience
in Adolescence course, May 3-9.

Cosme, D. (2020) Facilitating widespread behavior change: The role of
self and social relevance information sharing and virality. Guest
lecture presented at the Bard College PSY334 Science of Goal Pursuit
course, December 10, 2020.

Cosme, D. (2015) Emotion and the brain. Guest lecture presented at the
University of Oregon SCAN354 Mind Games: Emotion and Subjectivity in
Nordic Prose course, January 15.
:::

:::

## Informal teaching experience

<b>Teaching experiences in informal settings, such as meetings, clubs, and workshops</b><br>

::: grid

::: {.g-col-12 .g-col-md-12}
**Winter 2018 - Fall 2019 | Undergraduate Research Methods Meeting**  
Bi-weekly meeting for the undergraduate research assistants I supervised

**[Spring 2018 | Open Neuroscience Workshop](https://github.com/brainhack-eugene/open_neuroscience_workshop)**  
1-day workshop for University of Oregon students, staff, and community
members

**[Fall 2017 | Modeling Developmental Change Workshop](https://osf.io/xuyjm)**  
2-day workshop for developmental cognitive neuroscientists attending the
Flux congress

**[Fall 2016 - Spring 2018 | Data Science Club](https://github.com/uodatascience)**  
Weekly meeting for graduate students at the University of Oregon
:::

:::

## Mentorship & supervision

<b>Trainees that I have mentored and supervised; notes indicate mentorship as part of a specific fellowship, program, or in a particular capacity.</b><br>

### Postdocs

```{r}
#| echo: false
#| message: false
#| warning: false
#| tbl-colwidths: [50, 10, 10, 30]

library(tidyverse)

mentees = read.csv("mentorship.csv")

mentees %>%
  filter(Stage == "Postdoc") %>%
  select(-Stage) %>%
  knitr::kable()
  
```

### Graduate students

```{r}
#| echo: false
#| message: false
#| warning: false
#| tbl-colwidths: [50, 10, 10, 30]

mentees %>%
  filter(Stage == "Graduate") %>%
  select(-Stage) %>%
  knitr::kable()
  
```

### Postbacs

```{r}
#| echo: false
#| message: false
#| warning: false
#| tbl-colwidths: [50, 10, 10, 30]

mentees %>%
  filter(Stage == "Postbac") %>%
  select(-Stage) %>%
  knitr::kable()
  
```

### Undergraduates

```{r}
#| echo: false
#| message: false
#| warning: false
#| tbl-colwidths: [50, 10, 10, 30]

mentees %>%
  filter(Stage == "Undergraduate") %>%
  select(-Stage) %>%
  knitr::kable()
  
```
